Supplemental Materials, Methods and Results

### 2021-04-28 ms.Rmd
Adding code to implement RVE across the board including pairwise comparisons

p_value <- function(x){
  if(x <= 0.0001) {tmp = "< 0.0001"}
  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
  if(x >= 0.01) {tmp =x}
  return(tmp)
}

@@

Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package (). Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables. For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects. For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or snake), population of origin and their interaction as fixed effects. Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii. Model interactions were first tesed using a Type III sum of squares ANOVA with a Saitterwaite degrees of freedom correction, and if found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model. For all individuals, we measured signals for multiple body regions, each of which were contratsed between multiple background measurements (i.e., 2 random leaves, and 2 random stems). For our main analyses we average estimates across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results). Deriving multiple contrats for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely indpeendent of each other. To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) on our final model to correct standard errors for within-individual non-independence. In all cases, our results and conclusions were not impacted (see supplmentary code).

>>>>

p_value <- function(x){
  if(x <= 0.0001) {tmp = "< 0.0001"}
  if(x <= 0.001 & x >= 0.0001) {tmp ="< 0.001"}
  if(x <= 0.01 & x >= 0.001) {tmp ="< 0.01"}
  if(x >= 0.01) {tmp = round(x, digits =2)}
  return(tmp)
}

@@

Chromatic and achromatic contrast from visual models, measured in just noticable differences (JNDs), were analysed using linear mixed effects models in R using the `lme4` package (). Across the different contexts (i.e., social contexts, predation and background environments), chromatic and acromatic contrasts were analysed in seperate models as response variables. For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects. For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or snake), population of origin and their interaction as fixed effects. Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were obersved on a native Kenyan background or their current introduced habitat background in Hawaii. Model interactions were first tesed using a Type III sum of squares ANOVA with a Saitterwaite degrees of freedom correction, and if found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model. For all individuals, we measured signals for multiple body regions, each of which were contratsed between multiple background measurements (i.e., top and bottoms of two leaves and two stems). For our main analyses we average estimates across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see Supplementary Results). Deriving multiple contrats for each individual body region and background resulted in a substantial number of comparisons, each of which were not completely indpeendent of each other. To ensure that non-indpendence of data did not impact our inferences we used a Robust Variance Estimator (RVE) on our final model to correct standard errors for within-individual non-independence. In all cases, our results and conclusions were not impacted (see supplmentary code).

### 2021-04-28 supp.Rmd
Adding code to implement RVE across the board including pairwise comparisons

---
title: "Supplemental Materials, Methods and Results"
author: Whiting et al. 
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: science.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: true
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
  ## and rounded to 2 digits      ## and rounded to 2 digits
  options(digits = 2)
```

@@

#########################################################
# Male-male Competition Displays
#########################################################
#Load data and clean a bit
Display_1 <- read.table("./data/Display_1.csv", 
                          header=TRUE, sep=",", dec=".", strip.white=TRUE)

#exclude bottom flank.
Display_1 <- Display_1 %>% 
              filter(!BodyRegion == "botflank") %>% as.data.frame()

# create as factor
    Display_1$BodyRegion <- factor(Display_1$BodyRegion)

@@

# Fit the model that esatimtes changes across body regions
ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Background 
                         + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)

# Fit model for linear hypothesis tests
Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
ConspicDisplayAC2 <- lmer(dL ~ PopBRInt + (1|ID), data=Display_1)

result1 = glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
                                                            "Hawaii.midflank - Kenya.midflank = 0",     
                                                            "Hawaii.topflank - Kenya.topflank = 0", 
                                                             "Hawaii.tailbase - Kenya.tailbase = 0" )))

## Display - Chromatic contrast (CC or dS)
## just this next section has Drew's code, which averages the backround (Leaf_B1, Leaf_B2, Leaf_T1, Leaf_T2, Stem1, Stem2) for each lizard body region measurement.
## This is the full model; however, we do not need background because the way the data is set up R is taking an average. See next model (background dropped)

# Fit the model that esatimtes changes across body regions
  ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Background 
                         + Population:Background + Population:BodyRegion + (1|ID), data=Display_1)
Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)

>>>>

---
title: "Supplemental Materials, Methods and Results"
author: Whiting et al. 
date: "`r Sys.Date()`"
bibliography: refs.bib
csl: science.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
## numbers >= 10^5 will be denoted in scientific notation,      ## numbers >= 10^5 will be denoted in scientific notation,
  ## and rounded to 2 digits      ## and rounded to 2 digits
  options(digits = 2)
```

@@

#########################################################
# Male-male Competition Displays
#########################################################
  # Load data and clean a bit
  Display_1 <- read.table("./data/Display_1.csv", 
                            header=TRUE, sep=",", dec=".", strip.white=TRUE)

  # Exclude bottom flank.
  Display_1 <- Display_1 %>% 
                filter(!BodyRegion == "botflank") %>% as.data.frame()

  # Create as factor
      Display_1$BodyRegion <- factor(Display_1$BodyRegion)

@@

##########
## Achromatic Contrast - dL
##########

    # Fit the model that esatimtes changes across body regions
    ConspicDisplayAC <- lmer(dL ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
    summary(ConspicDisplayAC)

    # Robust variance estimators (RVE). Mostly all consistent
    Display_subset <- Display_1 %>% filter(!is.na(sc_SVL), !is.na(Population), !is.na(BodyRegion))
    robust_ConspicDisplayAC <- coef_test(ConspicDisplayAC, cluster = Display_subset$ID, vcov="CR2")

    # ANOVA. 
        anova(Display_subset)

    # Fit model for linear hypothesis tests
           Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
            ConspicDisplayAC2 <- lmer(dL ~ -1 + sc_SVL + PopBRInt + (1|ID), data=Display_1)
            summary(ConspicDisplayAC2)
    # Fit RVE
     robust_ConspicDisplayAC2 <- coef_test(ConspicDisplayAC2, cluster = Display_subset$ID, vcov="CR2")

     # Contrasts hypothesis tests, but use RVE 
    result1 <-Wald_test(ConspicDisplayAC2, constraints = constrain_pairwise(c(2:9)), vcov="CR2", tidy = TRUE)
    result1$p_val_bonfer <- sapply(p.adjust(result1$p_val, "bonferroni"), function(x) p_value(x))
    result1 = glht(ConspicDisplayAC2, linfct = mcp(PopBRInt = c("Hawaii.gular - Kenya.gular = 0",
                                                                "Hawaii.midflank - Kenya.midflank = 0",     
                                                                "Hawaii.topflank - Kenya.topflank = 0", 
                                                                 "Hawaii.tailbase - Kenya.tailbase = 0" )))

##########
## Chromatic Contrast - dL
##########

# Fit the model that esatimtes changes across body regions
  ConspicDisplayCC <- lmer(dS ~ sc_SVL + Population + BodyRegion + Population:BodyRegion + (1|ID), data=Display_1)
Display_1$PopBRInt <- interaction(Display_1$Population,Display_1$BodyRegion)
